A Reduced Order Model for Porous Media Flows: Coupling Linear Regression with Principal Component Analysis (PCA) using High-Order Singular Value Decomposition (HOSVD) in Upscaling Calculations for Permeability in Reservoir Simulations

### 2018-12-03 Papers/AppliedMaths2018/Resubmission/Stochastic_Upscaling_UK_AMM_2018b.tex
files to be modified in the next 250min

Immiscible multi-fluid flows in porous media may be represented by the extended Darcy equation,
\begin{equation}\label{e:darcy_eqn}
  \mathbf{q}_{\alpha} = \mathbf{u}_{\alpha}S_\alpha=
  -\frac{\mathcal{K}_{{r}_\alpha}\mathbf{K}}{\mu_{\alpha}}\left(
  \nabla p_{\alpha} - {\mathbf{s}_{u}}_{\alpha} \right),
\end{equation}
where $\alpha$ designates a phase, $\mathbf{q}_{\alpha}$ is the $\alpha$-th phase Darcy velocity and $\mathbf{u}_{\alpha}$ is the saturation-weighted Darcy velocity. $\mu_{\alpha}$, $p_{\alpha}$, $\rho_{\alpha}$, and $\mathbf{s}_{{u}_\alpha}$ are phase dynamic viscosity, pressure, density and source term (which may include gravity and/or capillarity), respectively. $\mathcal{K}_{{r}_\alpha}\left(S_{\alpha}\right)$ is the phase relative permeability, which is a function of the phase saturation $S_{\alpha}\left(\mathbf{r},t\right)$, which in turn is a function of the vector position $\mathbf{r}$, and time, $t$.

In addition to the extended Darcy equation (Eqn.~\ref{e:darcy_eqn}), saturation conservative equations,
\begin{equation}
  \phi\displaystyle\frac{\partial S_{\alpha} }{\partial t} + \nabla
  \cdot \left( {\mathbf u}_{\alpha} S_{\alpha}\right) =
  s_{cty,\alpha},
  \label{saturation_equation}
\end{equation}
are discretised in space with CV basis functions, and in time with the $\theta$-method \cite{pavlidis_2016}. In Eqn.~\ref{saturation_equation}, $\phi$ is the porosity and $s_{cty}$ is a source term.

\medskip
The discretised saturation and Darcy equations are solved using a multigrid-like approach as described by \citeauthor{pavlidis2016}~\cite{pavlidis2016}. The numerical formulation is fully described by \citeauthor{Gomes_2017}~\cite{Gomes_2017} (see also \cite{adam_2016, salinas2015}). These numerical methods are embedded in the next-generation flow simulator Fluidity/IC-FERST model software\footnote{\href{http://multifluids.github.io}{http://multifluids.github.io}} (a full description of the model can be found in \cite{Gomes_2017, jackson_2013}). Although, the numerical formulation used here is relatively new, it is not the focus of this work. The result and discussion section (Section \ref{section:results_discussion}) refers to \citet{dawe_2008} for validation of the numerical formulation with respect to the 2D simulation. However, more extensive validation tests can be found in the aforementioned references \cite{adam_2016, Gomes_2017,jackson_2013, pavlidis2016, salinas2015}.

>>>>

Immiscible multi-fluid flows in porous media may be represented by the extended Darcy equation,
\begin{equation}\label{e:darcy_eqn}
  \mathbf{q}_{\alpha} = \mathbf{u}_{\alpha}S_\alpha=
  -\frac{\mathcal{K}_{{r}_\alpha}\mathbf{K}}{\mu_{\alpha}}\left(
  \nabla p_{\alpha} - {\mathbf{s}_{u}}_{\alpha} \right),
\end{equation}
where $\alpha$ designates a phase, $\mathbf{q}_{\alpha}$ is the $\alpha$-th phase Darcy velocity and $\mathbf{u}_{\alpha}$ is the saturation-weighted Darcy velocity. $\mu_{\alpha}$, $p_{\alpha}$, $\rho_{\alpha}$, and $\mathbf{s}_{{u}_\alpha}$ are phase dynamic viscosity, pressure, density and source term (which may include gravity and/or capillarity), respectively. $\mathbf{K}\left(\mathbf{r}\right)$ and $\mathcal{K}_{{r}_\alpha}\left(S_{\alpha}\right)$ are absolute and phase-relative permeability, respectively. The later is a function of the phase saturation $S_{\alpha}\left(\mathbf{r},t\right)$, which in turn is spatial- and time-dependent.

In addition to the extended Darcy equation (Eqn.~\ref{e:darcy_eqn}), saturation conservative equations,
\begin{equation}
  \phi\displaystyle\frac{\partial S_{\alpha} }{\partial t} + \nabla
  \cdot \left( {\mathbf u}_{\alpha} S_{\alpha}\right) =
  s_{cty,\alpha},
  \label{saturation_equation}
\end{equation}
are discretised in space with CV basis functions, and in time with the $\theta$-method \cite{gomes_book_2012}. In Eqn.~\ref{saturation_equation}, $\phi$ is the porosity and $s_{cty}$ is a source term.

\medskip
The discretised saturation and Darcy equations are solved using a multigrid-like approach as described by \citeauthor{pavlidis2016}~\cite{pavlidis2016}. The numerical formulation is fully described by \citeauthor{Gomes_2017}~\cite{Gomes_2017} (see also \cite{adam_2016, salinas2015}). These numerical methods are embedded in the next-generation flow simulator Fluidity/IC-FERST model software\footnote{\href{http://multifluids.github.io}{http://multifluids.github.io}} (a full description of the model can be found in \cite{Gomes_2017, jackson_2013,salinas_2018}). Although, the numerical formulation used here is relatively new, it is not the focus of this work. The result and discussion section (Section \ref{section:results_discussion}) refers to \citet{dawe_2008} for validation of the numerical formulation with respect to the 2D simulation. However, more extensive validation tests can be found in the aforementioned references \cite{adam_2016, Gomes_2017,jackson_2013, pavlidis2016, salinas2015}.
